---
title: "DataCamp Class Assignment"
subtitle: "Hannah Korte, worked with Ollie Murphy"
author: "Hannah Korte"
date: "9/11/2021"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
a= "Hannah"
b1= 45.6
b2= "45.6"
c1= 0:3
```
Q1: What type of data is contained in the variable a

- Character

Q2: What type of data is contained in the variable b1

- Numeric

Q3: What type of data is contained in the variable b2

- Character

Q4: What happens when you add b1 and b2 and why?

- Error in b1 + b2 : non-numeric argument to binary operator

Q5: Are the variables b1 and c1 of the same type?

- No

Q6: Explain what happens when you add b1 and c1. Consider both the number of elements in each variable and the data types.

- R adds the numeric value of b1 (45.6) to each number in the c1 vector, the results are 45.6 46.6 47.6 48.6.

Q7: Show the R code you used to create v1.
```{r echo=TRUE}
v1= c(-2:2)
v1
```
Q8: Show the R code you used to create v2.
```{r echo=TRUE}
v2= v1*3
v2
```
Q9: Show the R code you used to calculate the sum of elements in v2.
```{r}
sum(v2)
```
```{r}
vec_4= 1:12
vec_4
```
Q10: Show the code you used to create mat_1
```{r}
mat_1= matrix(vec_4, nrow= 3, ncol=4, byrow=TRUE)
mat_1
```
Q11: Show the code you used to create mat_2
```{r}
mat_2= matrix(vec_4, nrow=3, ncol=4,byrow=FALSE)
mat_2
```
Q12: Show the R code you used to create my_list_1.
```{r}
my_list_1=list("two"= 5.2,"one"="five point two","three"= c(0:5))
```
Q13: Show the R code that would select third element of the list.
```{r}
my_list_1[3]
```
Q14: Show the R code that selects the list element with the name “one”.
```{r}
my_list_1$one
```
Q15: Show the R code that you used to create my_bool_vec.
```{r}
my_vec= rep(1:3,5)
my_bool_vec=c(my_vec==3)
data.frame(my_vec,my_bool_vec)
```
Q16: Show the R code that you used to perform the subsetting.
```{r}
my_vec[my_bool_vec]
```

